\part{Linear Systems}
\setcounter{section}{0}

\section{Theory}
\Def The Condition Number of a matrix is $$\operatorname{cond}(\mathbf{A}):=\norm{\MA}\norm{\MA^{-1}}$$ and is a measure for stability ($\approx 1$ for stability).
Note that it is infinite for singular matrices.
Intuitively $cond(A)$ catches the maximum stretching factor divided by the minimum stretching factor.
\Lemma If $m=n$ and $\mathbf{A}$ is regular/invertible, then its 2-norm condition number is  $\operatorname{cond}_{2}(\mathbf{A})=\sigma_{1} / \sigma_{n}$
\sep

\section{LU Decomposition}

There are two methods to choose pivots in the elimination process with a Permutation Matrix:
\begin{itemize}
	\item Partial Pivoting: Max of column
	\item Full Pivoting: Max of remaining matrix
\end{itemize}
\sep

\textbf{Complexities:}
\begin{itemize}
	\item Gauss elim. $\BigO(n^3)$, back substitution $\BigO(n^2)$
  	\item LU: decomp. $\BigO(n^3)$, solve $\BigO(n^2)$
  	\item Inverse: compute $\BigO(n^3)$, solve $\BigO(n^2)$ 
\end{itemize}

\section{Exploiting Structure when Solving Linear Systems}

\Theorem[Low Rank Modification] Use the \textit{Sherman-Morrison-Woodbury}
Formula to compute the inverse of slightly modified (rank-1-modification) matrix.
$$
{\scriptstyle \left(\mathbf{A}+\mathbf{U V}^{H}\right)^{-1}=\mathbf{A}^{-1}-\mathbf{A}^{-1} \mathbf{U}\left(\mathbf{I}+\mathbf{V}^{H} \mathbf{A}^{-1} \mathbf{U}\right)^{-1} \mathbf{V}^{H} \mathbf{A}^{-1}
}$$

\section{Matrix Storage Formats}
\textbf{Dense Matrices} are either stored in Column-Major Ordering or Row-Major Ordering. By default Eigen employs Column-Major Ordering.
\sep
\textbf{Sparse Matrices} can be stored in
\begin{itemize}
	\item COO/Triplet: Vector of triplets of (row, col, value)
  	\item CCS: Compressed Column Storage
  	\item CRS: Compressed Row Storage
\end{itemize}
where CRS consists of the arrays:
\begin{itemize}
	\item val: All non-zero values in Row-Major Order
	\item col\_idx: Column index of each entry in val
  	\item row\_ptr: Pointers to first elements of row
\end{itemize}

Note that $\text{size}(row\_ptr)=m+1$ (Sentinel).


% Definition von ODE
%\sep
%\Recipe[Householder QR]